---
title: "GammaSqRoot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(grid)
library(gridExtra)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
devtools::install_github("simonecollier/lizardHMM")
library(lizardHMM)
library(tinytex)
library(forecast)
```

## Retrieve and clean Data

```{r, include = TRUE}
library(readxl)
Pm_33_20Jul2021_1_ <- read_excel("Pm_33_20Jul2021 (1).xlsx")
trial1 <- Pm_33_20Jul2021_1_
trial1 <- subset (trial1, select = -c(frame, ...5))
```

Reduce the values to one frame per second
```{r}
skip_xpos <- as.numeric(trial1$pos_x[seq(28, length(trial1$pos_x), 29.97)])
skip_ypos <- as.numeric(trial1$pos_y[seq(28, length(trial1$pos_y), 29.97)])
steps <- euclidean_distance(skip_xpos, skip_ypos)
stepdata <- data.frame(step_length = steps,
                       time = seq(from = 1, by = 1, #29.97fps/29.97 -> 1fps
                                  length.out = length(steps)))
num_time <- nrow(stepdata)

#filter out data in sqrt_steps that is +/- inf
stepdata$sqrt_steps <- sqrt(stepdata$step_length)
inf_ind <- which(is.infinite(stepdata$sqrt_steps))
for (i in inf_ind) {
  stepdata$sqrt_steps[i] <- NA
}

#filter out data resulting from the tracking dot switching back and forth: indices of stepdata that have a step_length greater than 0.4
ind <- which(stepdata$log_steps[(1):(6900)] >2) 
for(i in ind) {
  stepdata$log_steps[i] <- NA
}
```

Plot 2 hour timeseries

```{r}
ggplot(data = stepdata, aes(x=time, y = sqrt_steps)) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = ggplot2::element_blank(),
        panel.grid.minor = ggplot2::element_blank(),
        plot.title = ggplot2:: element_blank()) +
  labs(x = 'Time (s)', y = 'Sqrt(Step Length)')
```
Histogram of logged step lengths:

```{r}
ggplot(data = stepdata, aes(x = sqrt_steps)) +
  geom_histogram(binwidth = 0.01,
                 colour = "cornsilk4",
                 fill = "white") +
  theme_bw() + 
  theme(panel.grid.major = ggplot2::element_blank(),
        panel.grid.minor = ggplot2::element_blank(),
        plot.title = ggplot2::element_blank()) +
  labs(x = 'sqrt(Step Length) per 0.2 s', y = '') + xlim(c(0,2)) + ylim(c(0,150))
```

Define Parameters for Gamma Distribution

```{r}
num_variables = 1
num_subjects = 1
num_covariates = 0
x             <- array(dim = c(num_time, num_variables, num_subjects))
x[, 1, 1]     <- stepdata$sqrt_steps
# no covariates so the design matrix is just a column of ones
design           <- list(matrix(0, nrow = num_time, ncol = 1))
design[[1]][, 1] <- 1
```

Fit sqrt of step lengths with a 2-state gamma HMM:

```{r setup, include=FALSE}
zweight0 <- list(c(0.4))
beta0 <- matrix(c(-1,
                  -1), ncol = 1, nrow = 2, byrow = TRUE)
delta0 <- list(c(1/2, 1/2))
#Package for parallel computations
library(parallel)

#Create cluster of size ncores
ncores <- detectCores() - 1
cl <- makeCluster(getOption("cl.cores", ncores))
#Export objects needed in parallelised function to cluster
clusterExport(cl, list("x", "gam0_fit_hmm", "beta0", "delta0", "design", "zweight0"))

num_states <- 2
#set.seed(12902)
niter <- 25
allM <- list()

#Create list of starting values
allPar0 <- lapply(as.list(1:niter), function(z) {
  #Step Length Mean
  alpha0 <- list(matrix(runif(2, min = c(.25, .5), max=c(.5, 1.5)), 
                     ncol = 2, nrow = 1, byrow = TRUE))
  #Step length SD
  theta0 <- list(matrix(runif(2, min=c(.5, 2), max=c(2, 4)), 
                        ncol = 2, nrow = 1, byrow = TRUE))
  
  
  return(list(alpha = alpha0, theta = theta0))
})

#mu0 = allPar0[]
#sigma0 = allPar0[1][1]

allm_parallel <- parLapply(cl = cl, X = allPar0, fun = function(par0){
  
  m <- gam0_fit_hmm(x, design, num_states = 2, num_variables = 1, num_subjects =1,num_covariates = 0, alpha0 = par0$alpha, theta =par0$theta, zweight0, beta0 = beta0, delta0, iterlim = 200, hessian = TRUE)
  
  return(m)
})

View(allm_parallel)
mllks <- numeric()
for (i in 1:niter) {
  mllks <- c(mllks, allm_parallel[[i]]$max_loglikelihood)
}


#Extract likelihoods of fitted models
#mllks <- unlist(lapply(allM, function(m) m$hmm$max_loglikelihood))

whichbest <- which.min(mllks)
whichbest
sqstep_fit2 <- allm_parallel[[16]]
sqstep_fit2 
```

```{r, include = TRUE}
CI <- gam0_ci(sqstep_fit2, state_dep_dist_pooled = FALSE, n = 100, level = 0.975)
CI
```

```{r}
viterbi <- gam0_viterbi(x, sqstep_fit2)
timeseries_plot(x, viterbi, num_subjects, num_variables, end = 900)
```


Histogram of sqrt(step lengths) with state dependent Gamma distributions & CIs
```{r}
gam0_hist_ci(x, viterbi, num_states, num_subjects, num_variables, sqstep_fit2, 
             state_dep_dist_pooled = FALSE, width = 0.01, n = 100, level = 0.975,
             x_step = 0.01, xaxis = list(list(c(0,10))))
#whats a situation where state_dep_dist_pooled  = TRUE
```


Pseudo Residuals

```{r}
sqstep_psr2 <- gam0_forecast_psr(x, sqstep_fit2)
psr_plot(sqstep_fit2, num_subjects)
psr_hist(sqstep_fit2, num_subjects)
```

```{r}
psr_qq(sqstep_fit2, 1)
```

```{r}
sqstep_psr2 <- gam0_forecast_psr(x, sqstep_fit2)
psr_acf(sqstep_psr2, 1)
```

Fit sqrt of step lengths with a 3-state gamma HMM:

```{r setup, include=FALSE}
zweight0 <- list(c(0.4))
beta0 <- matrix(c(-1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1), ncol = 1, nrow = 6, byrow = TRUE)
delta0 <- list(c(1/3, 1/3, 1/3))
#Package for parallel computations
library(parallel)

#Create cluster of size ncores
ncores <- detectCores() - 1
cl <- makeCluster(getOption("cl.cores", ncores))
#Export objects needed in parallelised function to cluster
clusterExport(cl, list("x", "gam0_fit_hmm", "beta0", "delta0", "design", "zweight0"))

num_states <- 3
#set.seed(12902)
niter <- 25
#allM <- list()

#Create list of starting values
allPar0 <- lapply(as.list(1:niter), function(z) {
  #Step Length Mean
  alpha0 <- list(matrix(runif(3, min = c(.5, 1, 2), max=c(2, 3, 4)), 
                     ncol = 3, nrow = 1, byrow = TRUE))
  #Step length SD
  theta0 <- list(matrix(runif(3, min=c(.5, .75, 1), max=c(1, 1.5, 2)), 
                        ncol = 3, nrow = 1, byrow = TRUE))
  
  
  return(list(alpha = alpha0, theta = theta0))
})

#mu0 = allPar0[]
#sigma0 = allPar0[1][1]

allm_parallel <- parLapply(cl = cl, X = allPar0, fun = function(par0){
  
  m <- gam0_fit_hmm(x, design, num_states = 3, num_variables = 1, num_subjects =1,num_covariates = 0, alpha0 = par0$alpha, theta =par0$theta, zweight0, beta0 = beta0, delta0, iterlim = 200, hessian = TRUE)
  
  return(m)
})

mllks <- numeric()
for (i in 1:niter) {
  mllks <- c(mllks, allm_parallel[[i]]$max_loglikelihood)
}


#Extract likelihoods of fitted models
#mllks <- unlist(lapply(allM, function(m) m$hmm$max_loglikelihood))

whichbest <- which.min(mllks)
whichbest
sqstep_fit3 <- allm_parallel[[19]]
sqstep_fit3
```

```{r, include = TRUE}
CI <- gam0_ci(sqstep_fit3, state_dep_dist_pooled = FALSE, n = 100, level = 0.975)
CI
```
```{r}
viterbi <- gam0_viterbi(x, sqstep_fit3)
timeseries_plot(x, viterbi, num_subjects, num_variables, end = 2100)
```

Histogram of sqrt(step lengths) with state dependent Gamma distributions & CIs
```{r}
gam0_hist_ci(x, viterbi, num_states =3, num_subjects, num_variables, sqstep_fit3, 
             state_dep_dist_pooled = FALSE, width = 0.01, n = 100, level = 0.975,
             x_step = 0.01, xaxis = list(list(c(0,10))))
#whats a situation where state_dep_dist_pooled  = TRUE
```

Fit sqrt of step lengths with a 4-state gamma HMM:

```{r setup, include=FALSE}
zweight0 <- list(c(0.4))
beta0 <- matrix(c(-1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1,
                  -1), ncol = 1, nrow = 12, byrow = TRUE)
delta0 <- list(c(1/4, 1/4, 1/4, 1/4))
#Package for parallel computations
library(parallel)

#Create cluster of size ncores
ncores <- detectCores() - 1
cl <- makeCluster(getOption("cl.cores", ncores))
#Export objects needed in parallelised function to cluster
clusterExport(cl, list("x", "gam0_fit_hmm", "beta0", "delta0", "design", "zweight0"))

num_states <- 4
#set.seed(12902)
niter <- 25
#allM <- list()

#Create list of starting values
allPar0 <- lapply(as.list(1:niter), function(z) {
  #Step Length Mean
  alpha0 <- list(matrix(runif(4, min = c(.5, 1, 2, 3), max=c(3, 4, 5, 6)), 
                     ncol = 4, nrow = 1, byrow = TRUE))
  #Step length SD
  theta0 <- list(matrix(runif(4, min=c(.25, .5, .75, 1), max=c(1, 1.5, 1.75, 2)), 
                        ncol = 4, nrow = 1, byrow = TRUE))
  
  
  return(list(alpha = alpha0, theta = theta0))
})

#mu0 = allPar0[]
#sigma0 = allPar0[1][1]

allm_parallel <- parLapply(cl = cl, X = allPar0, fun = function(par0){
  
  m <- gam0_fit_hmm(x, design, num_states = 4, num_variables = 1, num_subjects =1,num_covariates = 0, alpha0 = par0$alpha, theta =par0$theta, zweight0, beta0 = beta0, delta0, iterlim = 200, hessian = TRUE)
  
  return(m)
})

mllks <- numeric()
for (i in 1:niter) {
  mllks <- c(mllks, allm_parallel[[i]]$max_loglikelihood)
}


#Extract likelihoods of fitted models
#mllks <- unlist(lapply(allM, function(m) m$hmm$max_loglikelihood))

whichbest <- which.min(mllks)
whichbest
sqstep_fit4 <- allm_parallel[[1]]
sqstep_fit4
```

```{r, include = TRUE}
CI <- gam0_ci(sqstep_fit4, state_dep_dist_pooled = FALSE, n = 100, level = 0.975)
CI
```

```{r}
viterbi <- gam0_viterbi(x, sqstep_fit4)
timeseries_plot(x, viterbi, num_subjects, num_variables, end = 900)
```

